A spiral galaxy's mass distribution uncovered through lensing and dynamics

### 2015-11-16 archive/v3/docs/abstract_v3.01.tex
More corrections; abstract is now 250 words long; edited introduction.

We analyse the stellar and dark matter distribution in the spiral galaxy SDSS J1331+3628 (J1331) by means of two independent methods: gravitational lensing and dynamical modelling. Hubble Space Telescope (HST) imaging by the SWELLS survey shows, that J1331's bulge is superimposed by a quadruplet of extended lensing images. By fitting a gravitational potential model to the image positions, we tightly constrain the mass inside the Einstein radius $R_\text{ein}=(0.91\pm0.02)''(\simeq1.83\pm0.04~\text{kpc})$ to within 4\%: $M_\text{ein} = (7.8\pm0.3) \cdot 10^{10} M_\odot$. According to long-slit major axis stellar kinematics, J1331 has a counter-rotating stellar core inside $\sim 2''(\simeq4.02~\text{kpc})$. We model the observed stellar kinematics in J1331's central regions by finding Multi-Gaussian Expansion (MGE) models for the stellar and dark matter distribution that solve the axisymmetric Jeans equations. We find that J1331 requires a steep total mass-to-light ratio gradient in the center to reproduce the observed stellar kinematics. The best fit dynamical model predicts a total mass inside the Einstein radius consistent with the lens model, and vice versa the lens model gives an successful prediction for the observed kinematics in the galaxy center. For a dynamical model including a NFW dark matter halo,  we constrain the halo to have virial velocity $v_{200} \simeq 240 \pm 40$ km/s and a concentration of $c_{200} \simeq 8 \pm 2$ in case of a moderate tangential velocity anisotropy of $\beta_z \simeq -0.4 \pm 0.1$. The NFW halo models can successfully reproduce the signatures of J1331's counter-rotating stellar core and predict J1331's stellar and gas rotation curve at larger radii. However, all these models are more massive than expected from the stellar velocity dispersion at larger radii alone. We speculate that this could indicate a non-trivial re-distribution of matter due to a possible merger event in J1331's past. This could have resulted in a stellar core in J1331's very center with a Milky-Way-like Kroupa/Chabrier initial mass function (IMF), surrounded by a extended stellar bulge population with an IMF even more bottom-heavy than a Salpeter IMF \Wilma{[TO DO: Word limit is 250.]}.

>>>>

We analyse the stellar and dark matter distribution in the spiral galaxy SDSS J1331+3628 (J1331) by means of two independent methods: gravitational lensing and dynamical modelling. Hubble Space Telescope (HST) imaging by the SWELLS survey shows, that J1331's bulge is superimposed by a quadruplet of extended lensing images. By fitting a gravitational potential model to the image positions, we tightly constrain the mass inside the Einstein radius $R_\text{ein}=(0.91\pm0.02)''(\simeq1.83\pm0.04~\text{kpc})$ to within 4\%: $M_\text{ein} = (7.8\pm0.3) \cdot 10^{10} M_\odot$. According to long-slit major axis stellar kinematics, J1331 has a counter-rotating stellar core inside $\sim 2''(\simeq4.02~\text{kpc})$. We model the observed stellar kinematics in J1331's central regions by finding Multi-Gaussian Expansion (MGE) models for the stellar and dark matter distribution that solve the axisymmetric Jeans equations. We find that J1331 requires a steep total mass-to-light ratio gradient in the center to reproduce the observed stellar kinematics. The best fit dynamical model predicts a total mass inside the Einstein radius consistent with the lens model, and vice versa the lens model gives an successful prediction for the observed kinematics in the galaxy center. For a dynamical model including a NFW dark matter halo,  we constrain the halo to have virial velocity $v_{200} \simeq 240 \pm 40~\text{km s}^{-1}$ and a concentration of $c_{200} \simeq 8 \pm 2$ in case of a moderate tangential velocity anisotropy of $\beta_z \simeq -0.4 \pm 0.1$. The NFW halo models can successfully reproduce the signatures of J1331's counter-rotating stellar core and predict J1331's stellar and gas rotation curve at larger radii. However, all these models are more massive than expected from the stellar velocity dispersion at larger radii alone. We speculate that this could indicate a non-trivial re-distribution of matter due to a possible merger event in J1331's past. This could have resulted in a stellar core in J1331's very center with a Milky-Way-like Kroupa/Chabrier initial mass function (IMF), surrounded by a extended stellar bulge population with an IMF even more bottom-heavy than a Salpeter IMF \Wilma{[TO DO: Word limit is 250.]}.

### 2015-11-16 docs/data_v3.02.tex
More corrections; abstract is now 250 words long; edited introduction.

\paragraph{Stellar kinematics.} For the dynamical modelling we also use the stellar kinematics for J1331 measured by \citet{SWELLSV}. They obtained long-slit spectra along J1331's major-axis with the Low Resolution Imaging Spectrograph (LRIS) on the Keck I 10m telescope. The width of the slot was $1''$ and the seeing conditions had a FWHM of $\sim 1.1''$. Spectra for spatial bins of different widths along the major axis were extracted. They measured line-of-sight stellar rotation velocities ($v_\text{rot}$) and stellar velocity dispersion ($\sigma$) by fitting Gaussian line profiles to emission \Wilma{[TO DO: Check if really emission lines or, more likely, absorption lines]} lines in these spectra \Wilma{[TO DO: Find out which lines]}. Gas kinematics were extracted from fits to H$\alpha$ and NII lines, as tracers for ionized gas.
\\The stellar kinematics, $v_\text{rot}$, $\sigma$ and $v_\text{rms}^2=v_\text{rot}^2 + \sigma^2$ are shown in Figure \ref{fig:kinematics}. The rotation curve reveals a counter-rotating core within $2''\simeq$ 4 kpc. Outside of $\sim 3.5''$ there is a steep drop in the dispersion, which is expected at the boundary between the pressure supported bulge and the rotationally supported disk, which appears around this radius in the F450W filter in Figure \ref{fig:F450W}. However, in the brighter F814W filter in Figure \ref{fig:F814W}  the large reddish bulge extends out to $\sim5''$. 
\\Inside of $\sim 4''$, the data appears to be symmetric, outside of this the assumption of axisymmetry seems not to be valid anymore, considering the data. We add -2.3 km/s to the $v_\text{rot}$ to ensure $v_\text{rot}(R'=0) \sim 0$ as a possible correction term for a misjudgement of the systemic velocity. We also symmetrize the data within $4''$ and asign a minimum error of $\delta v_\text{rms} > 5$ km/s to the $v_\text{rms}$ data. In the JAM modelling, which is based on the assumption of axisymmetry, only stellar kinematics with $R'  \lesssim 2.5''$ or $R' \lesssim 4''$ are used. Another reason to restrict to modelling on the bulge region is that our MGE in Table \ref{tab:MGEF814W} is only a good representation of J1331's F814W light distribution inside $\sim 5''$.\\

>>>>

\paragraph{Stellar kinematics.} For the dynamical modelling we also use the stellar kinematics for J1331 measured by \citet{SWELLSV}. They obtained long-slit spectra along J1331's major-axis with the Low Resolution Imaging Spectrograph (LRIS) on the Keck I 10m telescope. The width of the slot was $1''$ and the seeing conditions had a FWHM of $\sim 1.1''$. Spectra for spatial bins of different widths along the major axis were extracted. They measured line-of-sight stellar rotation velocities ($v_\text{rot}$) and stellar velocity dispersion ($\sigma$) by fitting Gaussian line profiles to emission \Wilma{[TO DO: Check if really emission lines or, more likely, absorption lines]} lines in these spectra \Wilma{[TO DO: Find out which lines]}. Gas kinematics were extracted from fits to H$\alpha$ and NII lines, as tracers for ionized gas.
\\The stellar kinematics, $v_\text{rot}$, $\sigma$ and $v_\text{rms}^2=v_\text{rot}^2 + \sigma^2$ are shown in Figure \ref{fig:kinematics}. The rotation curve reveals a counter-rotating core within $2''\simeq$ 4 kpc. Outside of $\sim 3.5''$ there is a steep drop in the dispersion, which is expected at the boundary between the pressure supported bulge and the rotationally supported disk, which appears around this radius in the F450W filter in Figure \ref{fig:F450W}. However, in the brighter F814W filter in Figure \ref{fig:F814W}  the large reddish bulge extends out to $\sim5''$. 
\\Inside of $\sim 4''$, the data appears to be symmetric, outside of this the assumption of axisymmetry seems not to be valid anymore, considering the data. We add $-2.3~\text{km s}^{-1}$ to the $v_\text{rot}$ to ensure $v_\text{rot}(R'=0) \sim 0$ as a possible correction term for a misjudgement of the systemic velocity. We also symmetrize the data within $4''$ and asign a minimum error of $\delta v_\text{rms} > 5~\text{km s}^{-1}$ to the $v_\text{rms}$ data. In the JAM modelling, which is based on the assumption of axisymmetry, only stellar kinematics with $R'  \lesssim 2.5''$ or $R' \lesssim 4''$ are used. Another reason to restrict to modelling on the bulge region is that our MGE in Table \ref{tab:MGEF814W} is only a good representation of J1331's F814W light distribution inside $\sim 5''$.

 \Wilma{[TO DO: Maybe mention the following details better in the data section.... "SWELLS picked galaxies in SDSS whose spectra had two different redshifts within $3''$, which makes it likely to capture strong lenses with typical Einstein radii of $1''$. Sufficiently inclined galaxies were picked by eye. Follow-up high resolution imaging with the Hubble Space Telescope's (HST) Wide-Field Planetary Camera 2 (WFPC2) was performed to confirm their lensing status \citep{SWELLSI}."]}

 \Wilma{[TO DO: Include better in data section. "of approximate Hubble type Sb at right ascension = 202.91800$^\circ$ and a declination = 36.46999$^\circ$ (epoch J2000)"]}

### 2015-11-16 docs/discussion_ML_v3.01.tex
More corrections; abstract is now 250 words long; edited introduction.

We've presented different dynamical models of the central region of J1331, some of them capturing the observed kinematics well, but none of them work both at small and large radii. In the following we will discuss possible reasons for these deviations by also comparing our results to previous work.

@@

Further knowledge of the stellar populations within J1331 and their initial mass functions (IMF) would be useful. In particular a sophisticated guess for the \emph{stellar} mass-to-light ratio in J1331's bulge could be compared to our very reliable measurement of the \emph{total} mass-to-light ratio inside the Einstein radius (see Table \ref{tab:einsteinML} ) to support or contradict the presence of a lot of dark matter in the bulge. \\

Traditional choices for the IMF are the bottom-heavy IMF by \citet{Salpeter1955},
$$\xi(m) \propto m^{-x}, x=2.35,$$
(where $\xi(m) \diff m$ is the number of stars with mass $m$ in $[m,m+\diff m]$) and the IMFs by \citet{2002Sci...295...82K} and \citet{Chabrier2003}, which are in agreement with each other and predict less low-mass stars. \\

\citet{Ferreras} found a relation between the central stellar velocity dispersion $\sigma_0$ in early-type galaxies and the IMF slope. A higher $\sigma_0$ suggests a more bottom-heavy IMF. For an unimodal (Salpeter-like) IMF and a $\sigma_0 \propto 200 $ km/s for J1331 (see Figure \ref{fig:kinematics}) they predict $x \approx 2.33$ (see Figure 4 in \citet{Ferreras}), which is close to the standard Salpeter slope. These findings are supported by \citet{2014MNRAS.438.1483S}, who found for SDSS galaxies also a relation between unimodal (single power-law) IMF slope and stellar velocity dispersion. Their Table 2 lists $x(\sigma_*=190 \pm 20 \text{ km s}^{-1}) = 2.08 \pm 2.08$ and $x(\sigma_*= 230 \pm 20 \text{ km s}^{-1}) = 2.33 \pm 0.4$. When assuming a bi-modal (Kroupa-equivalent-like) IMF, \citet{Ferreras} predict $x \approx 2.85$ (see Figure 4 in \citet{Ferreras}) for J1331's central velocity dispersion. This is more bottom-heavy than the standard \citet{2002Sci...295...82K} IMF. Overall the central velocity dispersion suggests a rather bottom-heavy IMF J1331's bulge and therefore large stellar mass-to-light ratio.\\

\citet{SWELLSI} estimated J1331's stellar bulge mass given a Salpeter IMF and measured the I-band AB magnitude of the bulge. Transformed to a stellar I-band mass-to-light ratio (see Table \ref{tab:previousresults}) this would correspond to $\Upsilon_\text{I,*}^\text{sal} = 4.7 \pm 1.2$.\\

We compare our findings of the stellar mass-to-light ratios with the study by \cite{SWELLSV}: \cite{SWELLSV} found that the bulge of J1331 has an IMF more bottom-heavy than the Salpeter IMF from fitting a NFW halo to (1) the Einstein mass and (2) gas kinematics at larger radii $\gtrsim 8''$. In Section \ref{sec:results_JAM_NFW} we fitted the stellar kinematics within $\sim 3.5''$ and the Einstein mass with a NFW halo and the best fit $\Upsilon_\text{I,*} = 4.2 \pm 0.2$ (see Table \ref{tab:modelB4_bestfit}) indicates a less bottom-heavy IMF than the Salpeter IMF. \cite{SWELLSV}  found systematically lower NFW halos ($v_\text{circ}(5'') \sim 120 \text{km s}^{-1}$ according to their Figure 2) than we did ($v_\text{circ}(5'') \sim 200 \text{km s}^{-1}$, see Figure \ref{fig:kinematics}). We note however, that our results would be much closer to those by \citet{SWELLSV}, if we only used the lensing results at small radii as they did: The total mass-to-light ratio of $\Upsilon_\text{I,tot}^\text{ein} = 5.6$ inside the Einstein radius (see Table \ref{tab:einsteinML}). In case of a lower mass dark matter halo, the largest contribution to this mass-to-light ratio would then be stellar mass. And the corresponding IMF would be more bottom-heavy than Salpeter,  like what \citet{SWELLSV} found. The circular velocity curve $v_\text{circ}(R) = \left. \sqrt{R \frac{\partial \Phi}{\partial R}} \right|_{z=0}$ \citet{SWELLSV} find for J1331 peaks at $v_\text{circ}(R\sim0.9'') \sim 350 \text{km s}^{-1}$ (from lensing constraints only) and at $v_\text{circ}(R\sim0.9'') \sim 370 \text{km s}^{-1}$ (when including a dark matter halo and gas kinematics in the fit). We get a very similar circular velocity curve (at least at small radii) for a mass-follows-light model with constant $\Upsilon_\text{I,tot}^\text{ein} = 5.6$, which peaks at $v_\text{circ}(R\sim0.9'') \sim 360 \text{km s}^{-1}$. From this comparison follows, that the model by \citet{SWELLSV} corresponds approximately to a mass-follows-light model in the inner regions of J1331. In Section \ref{sec:results_JAM_SB} however we showed, that this is \emph{not} a good model for J1331 and won't be able to reproduce the central stellar $v_\text{rms}$ dip. On the other hand, because our NFW JAM models fail to predict the stellar kinematics at large radii, the true halo parameters are more likely closer to the findings by \citet{SWELLSV}.\\

All of this leaves the central dip unexplained: It cannot be due to tangential velocity anisotropy in the center alone (see Section \ref{sec:results_JAM_SB}, Figure \ref{fig:JAM_modelA2}). It also cannot be explained by a strong contribution of dark matter inside the bulge together with a moderate tagential velocity anisotropy, because the corresponding dark matter halos would still be too massive to fit the data in the outer regions (see Section \ref{sec:results_JAM_NFW}, Figure \ref{fig:modelB4_vrms}]).\\

An alternative JAM model was introduced in Section \ref{sec:results_JAM_SB}, where we used the surface brightness distribution without a halo component but together with an increasing mass-to-light ratio. We found that such a model would be perfectly consistent with the lensing results and predict a total mass-to-light ratio in the center of J1331 of  $\Upsilon_\text{I,tot}(R'\sim0) = 2.53$. This is very close to the stellar mass-to-light ratio we got from \citet{SWELLSI} estimates of the bulge's stellar mass using the Chabrier IMF $\Upsilon_\text{I,*}^\text{Chabrier} = 2.5 \pm 0.6$, i.e. less bottom-heavy (more bluish) population.\\

Another feature in the stellar kinematics, that none of our JAM models was able to reproduce in the slightest and which we therefore excluded in the modelling, was the dip in $v_\text{rms}$ around $\sim 6''$, which occurs around the transition from bulge to disk (see Figures \ref{fig:F814W} and \ref{fig:kinematics}).\\

The above discussion motivates the following speculation: In the absence of a strong dark matter contribution in the center, the overall stellar-mass-to-light ratio within the Einstein radius indicates a bottom-heavy IMF close to the Salpeter IMF, consistent with estimates from the velocity dispersion. At the same time the central dip can then be only explained, if there was an increase in stellar mass-to-light ratio with radius \textit{within} the bulge. \emph{If} there was a central stellar population with an IMF close to the \citet{Chabrier2003} IMF, surrounded by a more bottom-heavy population, the central stellar kinematics would be well explained and be fully consistent with the lensing results. (According to Figure \ref{fig:lenscompareboth} the lensing result might not predict a mass to light ratio gradient inside $R_\text{ein}$, but then again the mass slope $alpha=1$ was only weakly constrained.) The disk of J1331 has a lower $\Upsilon_\text{I,*}$ than the bulge (see Table \ref{tab:previousresults} according to \citet{SWELLSI}). Such a drop in $\Upsilon_\text{I,*}$ at the transition region from bulge to disk around $\sim 5''$ could lead to the observed drop in $v_\text{rms}$, while an increasing contribution of a lower mass dark matter halo at larger radii as found by \citet{SWELLSV}, would recover the kinematics in the out regions of J1331's disk.\\

>>>>

We've presented different dynamical models of the central region of J1331, some of them capturing the observed kinematics well, but none of them work both at small and large radii. In the following we will discuss possible reasons for these deviations, also by comparing our results to previous work.

@@

Further knowledge of the stellar populations within J1331 and their initial mass functions (IMF) would be useful. In particular a sophisticated guess for the \emph{stellar} mass-to-light ratio in J1331's bulge could be compared to our very reliable measurement of the \emph{total} mass-to-light ratio inside the Einstein radius (see Table \ref{tab:einsteinML} ) to support or contradict the presence of a significant amount of dark matter in the bulge.

Traditional choices for the IMF are the bottom-heavy IMF by \citet{Salpeter1955},
$$\xi(m) \propto m^{-x}, x=2.35,$$
where $\xi(m) \diff m$ is the number of stars with mass $m$ in $[m,m+\diff m]$, and the IMFs by \citet{2002Sci...295...82K} and \citet{Chabrier2003}, which are in agreement with each other and predict less low-mass stars.

\citet{Ferreras} found a relation between the central stellar velocity dispersion $\sigma_0$ in early-type galaxies and the IMF slope. A higher $\sigma_0$ suggests a more bottom-heavy IMF. For an unimodal (Salpeter-like) IMF and a $\sigma_0 \simeq 200~\text{km s}^{-1}$ for J1331 (see Figure \ref{fig:kinematics}) they predict $x \approx 2.33$, which is close to the standard Salpeter slope, also supported by \citet{2014MNRAS.438.1483S}. When assuming a bi-modal (Kroupa-equivalent-like) IMF, \citet{Ferreras} predict $x \approx 2.85$ for J1331's central velocity dispersion. This is more bottom-heavy than the standard \citet{2002Sci...295...82K} IMF. Overall the central velocity dispersion suggests a rather bottom-heavy IMF J1331's bulge and therefore large stellar mass-to-light ratio. \citet{SWELLSI} estimated J1331's stellar bulge mass given a Salpeter IMF and measured the I-band AB magnitude of the bulge. Transformed to a stellar I-band mass-to-light ratio (see Table \ref{tab:previousresults}) this would correspond to $\Upsilon_\text{I,*}^\text{sal} = 4.7 \pm 1.2$.

We compare our findings of the stellar mass-to-light ratios with the study by \cite{SWELLSV}: They found that the bulge of J1331 has an IMF more bottom-heavy than the Salpeter IMF from fitting a NFW halo to (1) the Einstein mass and (2) gas kinematics at larger radii $\gtrsim 8''$. In Section \ref{sec:results_JAM_NFW} we fitted the stellar kinematics within $\sim 3.5''$ and the Einstein mass with a NFW halo and the best fit $\Upsilon_\text{I,*} = 4.2 \pm 0.2$ (see Table \ref{tab:modelB4_bestfit}) indicates a less bottom-heavy IMF than the Salpeter IMF. \cite{SWELLSV}  found systematically lower NFW halos ($v_\text{circ}(5'') \sim 120~\text{km s}^{-1}$ according to their Figure 2) than we did ($v_\text{circ}(5'') \sim 200~\text{km s}^{-1}$, see Figure \ref{fig:kinematics}).

We note however, that our results would be much closer to those by \citet{SWELLSV}, if we only used the lensing results at small radii as they did: The total mass-to-light ratio of $\Upsilon_\text{I,tot}^\text{ein} = 5.6$ inside the Einstein radius (see Table \ref{tab:einsteinML}). In case of a lower mass dark matter halo, the largest contribution to this mass-to-light ratio would then be stellar mass, so that the corresponding IMF would be more bottom-heavy than Salpeter. The circular velocity curve that \citet{SWELLSV} find for J1331 peaks at $v_\text{circ}(R\sim0.9'') \sim 350~\text{km s}^{-1}$ (from lensing constraints only) and at $v_\text{circ}(R\sim0.9'') \sim 370~\text{km s}^{-1}$ (when including a dark matter halo and gas kinematics in the fit). We get a very similar circular velocity curve (at least at small radii) for a mass-follows-light model with constant $\Upsilon_\text{I,tot}^\text{ein} = 5.6$, which peaks at $v_\text{circ}(R\sim0.9'') \sim 360~\text{km s}^{-1}$. From this comparison follows, that the model by \citet{SWELLSV} corresponds approximately to a mass-follows-light model in the inner regions of J1331. In Section \ref{sec:results_JAM_SB} however we showed, that this is \emph{not} a good model for J1331 and won't be able to reproduce the central stellar $v_\text{rms}$ dip.

On the other hand, because our NFW JAM models fail to predict the stellar kinematics at large radii, the true halo parameters are more likely closer to the findings by \citet{SWELLSV}. All of this leaves the central dip unexplained: It cannot be due to tangential velocity anisotropy in the center alone (see Section \ref{sec:results_JAM_SB}, Figure \ref{fig:JAM_modelA2}). It also cannot be explained by a strong contribution of dark matter inside the bulge together with a moderate tagential velocity anisotropy, because the corresponding dark matter halos would still be too massive to fit the data in the outer regions (see Section \ref{sec:results_JAM_NFW}, Figure \ref{fig:modelB4_vrms}]).

An alternative JAM model was introduced in Section \ref{sec:results_JAM_SB}, where we used the surface brightness distribution without a halo component but together with an increasing mass-to-light ratio. We found that such a model would be perfectly consistent with the lensing results and predict a total mass-to-light ratio in the center of J1331 of  $\Upsilon_\text{I,tot}(R'\sim0) = 2.53$. This is very close to the stellar mass-to-light ratio we got from \citet{SWELLSI} estimates of the bulge's stellar mass using the Chabrier IMF $\Upsilon_\text{I,*}^\text{Chabrier} = 2.5 \pm 0.6$, i.e. less bottom-heavy (more bluish) population.

Another feature in the stellar kinematics, that none of our JAM models was able to reproduce in the slightest and which we therefore excluded in the modelling, was the dip in $v_\text{rms}$ around $\sim 6''$, which occurs around the transition from bulge to disk (see Figures \ref{fig:F814W} and \ref{fig:kinematics}).

The above discussion motivates the following speculation: In the absence of a strong dark matter contribution in the center, the overall stellar-mass-to-light ratio within the Einstein radius indicates a bottom-heavy IMF close to the Salpeter IMF, consistent with estimates from the velocity dispersion. At the same time the central dip can then be only explained, if there was an increase in stellar mass-to-light ratio with radius \textit{within} the bulge. \emph{If} there was a central stellar population with an IMF close to the \citet{Chabrier2003} IMF, surrounded by a more bottom-heavy population, the central stellar kinematics would be well explained and be fully consistent with the lensing results. (According to Figure \ref{fig:lenscompareboth} the lensing result might not predict a mass to light ratio gradient inside $R_\text{ein}$, but then again the mass slope $alpha=1$ was only weakly constrained.) The disk of J1331 has a lower $\Upsilon_\text{I,*}$ than the bulge (see Table \ref{tab:previousresults} according to \citet{SWELLSI}). Such a drop in $\Upsilon_\text{I,*}$ at the transition region from bulge to disk around $\sim 5''$ could lead to the observed drop in $v_\text{rms}$, while an increasing contribution of a lower mass dark matter halo at larger radii as found by \citet{SWELLSV}, would recover the kinematics in the out regions of J1331's disk.

### 2015-11-16 docs/discussion_merger_v3.01.tex
More corrections; abstract is now 250 words long; edited introduction.

\paragraph{Counter-rotating core and minor merger.} J1331 has a large counter-rotating stellar core within $\sim 2''$. This suggests a process in J1331's past in which two components with angular momenta oriented in opposite directions were involved. Accretion of gas on retrograde orbits and subsequent star formation could lead to a younger and counter-rotating stellar population \Wilma{[TO DO: REF]}. However, the to form enough stars such that the net rotation of the large and massive core is retrograde, a very large amount of gas would have had to be accreted by J1331- which is not very likely [TO DO: REF]. Another scenario are galaxy mergers. The products of major mergers of similarly massive disk galaxies are often quiescent, elliptical galaxies \citep{1972ApJ...178..623T} [TO DO: correct reference???]. Therefore a minor merger is more likely to have occured in J1331. The dense nucleus of a satellite galaxy on a retrograde orbit could survive the dissipationless accretion and could spiral to core due to tidal friction \citep{1984ApJ...287..577K}. Usually large bulges of spirals or ellipticals are reddest in the center and become bluer further out.  Because star formation in low-mass satellite galaxies occurred later than in massive galaxies \Wilma{[TO DO: REF]}, the merger remnant would now host a younger and bluer population with lower velocity dispersion within the older and redder bulge of the massive progenitor. Such a reverse color gradient could correspond to an increasing stellar mass-to-light ratio.
\\The minor merger theory could explain and support our findings from dynamical modelling and the previous sections, that J1331 might have an increasing stellar mass-to-light ratio within its core.

\paragraph{Possible modelling failures for merger remnants.} Another effect that minor mergers can have on galaxies are a misalignement (warp) of the photometric and kinematic major axis \Wilma{[TO DO: REF]}. Would this be the case for J1331 the assumption of axisymmetry is not true anymore and our dynamical modelling would fail.
\\Mergers also could have changed the 3D shape of the dark matter halo and the NFW halo was therefore not a good model for J1331's dark matter halo. We also tried to model J1331 with a cored logarithmic halo. However, due to degeneracies in the modelling, we were not able to either constrain the profile for a cored logarithmic halo, or to rule it out.

@@

Standard JAM modelling approaches seem not to work for J1331. A JAM model for J1331 would need to allow for stellar mass-to-light ratio gradients within the galaxy, velocity anisotropy and a dark matter halo. Because of degeneracies between stellar and dark mass, and matter distribution and anisotropy profile, such a dynamical model would not lead to very tight constraints on the model parameters.  A search for color gradients in J1331 and/or investigation of absorption line indices could support or contradict the suspicion of the existence of stellar mass-to-light ratio gradients in J133. Subsequently detailed stellar population analyses of the spectra taken along J1331's major axis should be conducted to constrain the mass-to-light ratio reliably. In addition, the dynamical modelling should use more of the available information on J1331 and fit dynamics (stellar and gas kinematics from \citet{SWELLSV}) simultaneously with the gravitational lensing (image positions, shape and even flux ratios) in a similar fashion to \citet{SWELLSIV}. To also model the extent, shape and flux of the lensing images, the method by \citet{2004ApJ...611..739T,2003ApJ...590..673W} could be employed, which models the surface brightness distribution of the images and source on a pixelated grid. However, for this to work a good model for the galactic extinction would be needed. High-resolution integral-field spectroscopy could help with this, allow for spatially resolved stellar population analysis and dynamical modelling in two dimensions. This would lead to much better understanding of J1331's structure and mass distribution and therefore answer questions on how minor mergers might modify spiral galaxies.

>>>>

\Wilma{[TO DO: Remove minor in merger.]}

\paragraph{Counter-rotating core and minor merger.} J1331 has a large counter-rotating stellar core within $\sim 2''$. This suggests a process in J1331's past in which two components with angular momenta oriented in opposite directions were involved. Accretion of gas on retrograde orbits and subsequent star formation could lead to a younger and counter-rotating stellar population \Wilma{[TO DO: REF]}. However, to form enough stars such that the net rotation of the large and massive core is retrograde, a very large amount of gas would have had to be accreted by J1331---which is not very likely [TO DO: REF].

Another scenario are galaxy mergers. Major mergers including large amounts of gas can form kinematically decoupled cores (KDCs), see e.g., \Wilma{[TO DO: REF: Lauren Hofman et al 2010 - read paper how this fits with the reverse color gradient]}. During a minor merger, the dense nucleus of a satellite galaxy on a retrograde orbit could survive the dissipationless accretion and could spiral to the core due to tidal friction \citep{1984ApJ...287..577K}. Usually large bulges of spirals or ellipticals are reddest in the center and become bluer further out.  Because star formation in low-mass satellite galaxies occurred later than in massive galaxies \Wilma{[TO DO: REF]}, the merger remnant would now host a younger and bluer population with lower velocity dispersion within the older and redder bulge of the massive progenitor. Such a reverse color gradient could correspond to an increasing stellar mass-to-light ratio.

The merger theory could explain and support our findings from dynamical modelling and the previous sections, that J1331 might have an increasing stellar mass-to-light ratio within its core.

\Wilma{[TO DO: Include in discussion: Kinematic twist due to merger. Bulge and disk have different kinematic major axis, and the measured velocities in the disk, which are off from the major axis, are much lower than the actual maximum vcirc. Could explain the dips in the outer regions.]}

\paragraph{Possible modelling failures for merger remnants.} Another effect that minor mergers can have on galaxies are a misalignement (warp) of the photometric and kinematic major axis \Wilma{[TO DO: REF]}. Would this be the case for J1331 the assumption of axisymmetry is not true anymore and our dynamical modelling would fail.

Mergers also could have changed the 3D shape of the dark matter halo and the NFW halo was therefore not a good model for J1331's dark matter halo. We also tried to model J1331 with a cored logarithmic halo. However, due to degeneracies in the modelling, we were not able to either constrain the profile for a cored logarithmic halo, or to rule it out.

\Wilma{[TO DO: Include in this discussion the findings from the color profile.]}

@@

Standard JAM modelling approaches seem not to work for J1331. A JAM model for J1331 would need to allow for stellar mass-to-light ratio gradients within the galaxy, velocity anisotropy and a dark matter halo. Because of degeneracies between stellar and dark mass, and matter distribution and anisotropy profile, such a dynamical model would not lead to very tight constraints on the model parameters.  A search for color gradients in J1331 and/or investigation of absorption line indices could support or contradict the suspicion of the existence of stellar mass-to-light ratio gradients in J133. Subsequently detailed stellar population analyses of the spectra taken along J1331's major axis should be conducted to constrain the mass-to-light ratio reliably.  \Wilma{[TO DO: Include in this discussion the findings from the color profile.]}

In addition, the dynamical modelling should use more of the available information on J1331 and fit dynamics (stellar and gas kinematics from \citet{SWELLSV}) simultaneously with the gravitational lensing (image positions, shape and even flux ratios) in a similar fashion to \citet{SWELLSIV}. To also model the extent, shape and flux of the lensing images, the method by \citet{2004ApJ...611..739T,2003ApJ...590..673W} could be employed, which models the surface brightness distribution of the images and source on a pixelated grid. However, for this to work a good model for the galactic extinction would be needed.

High-resolution integral-field spectroscopy could help with this, allow for spatially resolved stellar population analysis and dynamical modelling in two dimensions. \Wilma{[TO DO: Include in this discussion that this could also help test, if the vrms dip is maybe because of twist, i.e. the measurements where not done along the major axis in the disk]} This would lead to much better understanding of J1331's structure and mass distribution and therefore answer questions on how minor mergers might modify spiral galaxies.

\Wilma{[TO DO: Include in discussion: High resolution IFU: high spatial resolution --> to resolve small features. High spectral resolution --> because the velocity dispersion in the outer regions of the galaxy are very low and to measure them accurately, need high spectral resolution.]}
